IMF and Benchmark Forecasts

### 2023-09-15 manuscript/manuscript.Rmd
manuscript updates

# A short note on error handling

In almost all 72 cases, absolute error handling gives lower scores than directional error handling. The only exception is the inflation series for the IMF forecasts and horizon 0, where the expanding window and rolling window method give \textit{slightly} lower scores for the directional methodology. 
We thus decide to focus on the absolute errors in this document.

```{r loaddata, echo = FALSE}
scores <- fread(here("quantile_forecasts", "ci_scores.csv"))

wis_scores <- fread(here("quantile_forecasts", "ci_scores_avgcnt.csv")) |>
  data.table::copy() |>
  .d(, .(model, error_method, method, target, horizon, interval_score, dispersion, underprediction, overprediction)) |>
  setnames("model", "source")

crps_scores <- fread(here("quantile_forecasts", "sample_scores.csv"))

@@

@@

# Scores, by estimation method, Horizon and forecast source 
```{r, cpitable = FALSE}
#bring into long format
long_scores <- gather_scores(all_scores, "pcpi_pch")

@@

```{r overall_assessment, fig.dim=c(12,12), echo = FALSE}

scores_cvgshort <- fread(here("quantile_forecasts", "ci_scores_avgcnt.csv")) |>
  setnames("model", "source", skip_absent = TRUE)

point_scores <- data.table::fread(here("quantile_forecasts", "pointfc_scores.csv"))

    if(error_m == "directional"){
      plot_name <- "Directional Errors"
    } else {
      plot_name <- "Absolute Errors"
    }

    plt <- ggplot(decomp_data |> .d(error_method == error_m),
                          aes(x = source, 
                              fill = source,
                              y = score,
                              alpha = type)) +
      geom_bar(position="stack", stat="identity") +
      facet_wrap(~ horizon,
                 labeller = as_labeller(plot_horizon_label_short(4)),
                 nrow = 1) +
      #geom_hline(yintercept = 1) +
      scale_alpha_discrete(range = c(0.35,1),
                           labels = c("Overprediction",
                                      "Underprediction",
                                      "Dispersion"),
                           name = "Components of the WIS") +
      scale_x_discrete(name = "",
                       breaks = c("IMF", "bvar", "ar"),
                       labels = c("IMF", "BVAR", "AR")) +
      scale_y_continuous(name = "WIS", limits = c(0, maxval)) +
      scale_fill_brewer(palette = "Set2", 
                        name = "Model Type")+
      xlab("") +
      #labs(fill = "Model Type") +
      theme_uqimf() %+replace%
      theme(axis.text.x = element_text(size = 8, angle = 90, hjust = .5, vjust = .5, face = "plain"),
            strip.text = element_text(size = 8)) +
      ggtitle(plot_name) +
      theme(#legend.title=element_blank(),
            legend.position = "none")

  horizon_path_plot <- function(score_data,
                                score_rule,
                                ylab = score_rule, 
                                facet = NULL){

    path_plot <- ggplot(score_data, 
           aes(x = horizon,
               y = get(score_rule), 
               color = source)) +
      scale_color_brewer(palette = "Set2") +
      geom_line() +
      geom_point() +
      ylab(ylab) +
      xlab("Forecast Horizon") +
      guides(fill = "none", colour = "none") +
      theme_uqimf() %+replace%
      theme(legend.title=element_blank(),
            legend.position = "none")

  c50plot <- horizon_path_plot(cvgscores, "coverage_50", "CI Coverage - 50", facet = "error_method") +
    geom_hline(aes(yintercept = 0.5)) +
    scale_y_continuous(labels = paste0(seq(0, 100, by = 25), "%"),
                       breaks = seq(0,1, by = 0.25),
                       limits = c(0,1))

  c80plot <- horizon_path_plot(cvgscores, "coverage_80", "CI Coverage - 80", facet = "error_method") +
    geom_hline(aes(yintercept = 0.8)) +
    scale_y_continuous(labels = paste0(seq(0, 100, by = 25), "%"),
                       breaks = seq(0,1, by = 0.25),
                       limits = c(0,1))

  aeplot <- horizon_path_plot(point_scores |> .d(, error_method := "directional"), "meanae", "Absolute Error")

  squeplot <- horizon_path_plot(point_scores |> .d(, error_method := "directional"), "meansque", "Squared Error")

  crpsplot_rw <- horizon_path_plot(crps_scores_rw |>.d(, error_method := "directional"), "score", "CRPS by Sample")

  crpsplot_loo <- horizon_path_plot(crps_scores_loo |>.d(, error_method := "directional"), "score", "CRPS by Sample")

  ovr_plot <- 
    (dir_wis_plot | abs_wis_plot) /
    (plot_spacer()) /
    (plot_annotation(title = "hey")) /
    (c50plot | c80plot) /
    (plot_spacer()) /
    (squeplot | aeplot | crpsplot_rw | crpsplot_loo) +
    plot_layout(guides = "collect", 
                heights = c(0.2, 0.05, 0.2, 0.05, 0.2)) &
      plot_annotation(tag_levels = 'I')  &
      theme(legend.position = 'bottom', 
            legend.box="vertical", legend.margin=margin())

scores <- fread(here("quantile_forecasts", "ci_scores.csv"))
scores_cvgshort <- fread(here("quantile_forecasts", "cvg_pooled.csv"))

cvgdat_cpin <- data.table::fread(here("quantile_forecasts", "cvg_pooled_directionalsymmetric.csv")) |>
  setnames("model", "source") |>
  .d(, c( "target", "error_method", "method", "source", cols), with = FALSE) |>
  melt(id.vars = c("target", "error_method", "method", "source"),
       variable.name = "pilvl",
       value.name = "coverage") |>
  .d(, pilvl := gsub("^.*?coverage_","",pilvl)) |>
  .d(, pilvl := as.numeric(pilvl)/100) |>
  .d(target == "pcpi_pch") |>
  rbind(cvgdat_cpi)

cvgdat_gdpn <- data.table::fread(here("quantile_forecasts", "cvg_pooled_directionalsymmetric.csv")) |>
  setnames("model", "source") |>
  .d(, c( "target", "error_method", "method", "source", cols), with = FALSE) |>
  melt(id.vars = c("target", "error_method", "method", "source"),
       variable.name = "pilvl",
       value.name = "coverage") |>
  .d(, pilvl := gsub("^.*?coverage_","",pilvl)) |>
  .d(, pilvl := as.numeric(pilvl)/100) |>
  .d(target == "ngdp_rpch") |>
  rbind(cvgdat_gdp)

>>>>

```{r loaddata, echo = FALSE}
scores <- fread(here("scores", "ci_scores.csv"))

wis_scores <- fread(here("scores", "ci_scores_avgcnt.csv")) |>
  data.table::copy() |>
  .d(, .(model, error_method, method, target, horizon, interval_score, dispersion, underprediction, overprediction)) |>
  setnames("model", "source")

crps_scores <- fread(here("scores", "sample_scores.csv"))

@@

########

bvar_wis_scores <- fread(here("scores", "bvar_ci_scores_avgcnt.csv")) |>
  data.table::copy() |>
  .d(, .(model, target, horizon, interval_score, dispersion, underprediction, overprediction)) |>
  .d(,model := "bvar_qu") |>
  setnames("model", "source") |>
  .d(,error_method := "directional") |>
  .d(,method := "rolling window") #change later

bvar_wis_scores <- rbind(bvar_wis_scores, bvar_wis_scores |> copy() |> .d(, error_method := "absolute"))

@@

# Scores, by error method, Horizon and forecast source

```{r overall_assessment_ew, fig.dim=c(12,12), echo = FALSE, warning = FALSE}

scores_cvgshort <- fread(here("scores", "ci_scores_avgcnt.csv")) |>
  setnames("model", "source", skip_absent = TRUE)

bvar_scores_cvgshort <- fread(here("scores", "bvar_ci_scores_avgcnt.csv")) |>
  setnames("model", "source", skip_absent = TRUE) |>
  .d(, source := "bvar_qu") |>
  .d(, method := "rolling window") |>
  .d(, error_method := "absolute")

bvar_scores_cvgshort <- rbind(bvar_scores_cvgshort, bvar_scores_cvgshort |> copy() |> .d(, error_method := "directional"))

point_scores <- data.table::fread(here("scores", "pointfc_scores.csv"))

wis_scores <- rbind(wis_scores, bvar_wis_scores)
scores_cvgshort <- rbind(scores_cvgshort, bvar_scores_cvgshort)

    if(error_m == "directional"){
      plot_name <- "WIS - Directional Errors"
    } else {
      plot_name <- "WIS - Absolute Errors"
    }

    plt <- ggplot(decomp_data |> .d(error_method == error_m),
                          aes(x = source, 
                              fill = source,
                              y = score,
                              alpha = type)) +
      geom_bar(position="stack", stat="identity") +
      facet_wrap(~ horizon,
                 labeller = as_labeller(plot_horizon_label_short(4)),
                 nrow = 1) +
      #geom_hline(yintercept = 1) +
      scale_alpha_discrete(range = c(0.35,1),
                           labels = c("Overprediction",
                                      "Underprediction",
                                      "Dispersion"),
                           name = "Components of the WIS") +
      scale_x_discrete(name = "",
                       breaks = c("IMF", "bvar", "bvar_qu", "ar"),
                       labels = c("IMF", "BVAR", "BVAR_QU", "AR")) +
      scale_y_continuous(name = "WIS", limits = c(0, maxval)) +
      scale_fill_met_d("Hokusai1") +
      #scale_fill_brewer(palette = "Set2", 
      #                  name = "Model Type")+
      xlab("") +
      #labs(fill = "Model Type") +
      theme_uqimf() %+replace%
      theme(axis.text.x = element_text(size = 8, angle = 90, hjust = .5, vjust = .5, face = "plain"),
            strip.text = element_text(size = 8)) +
      ggtitle(plot_name) +
      theme(#legend.title=element_blank(),
            legend.position = "none")

  horizon_path_plot <- function(score_data,
                                score_rule,
                                ylab = score_rule, 
                                facet = NULL,
                                title = NULL){

    path_plot <- ggplot(score_data, 
           aes(x = horizon,
               y = get(score_rule), 
               color = source)) +

      scale_color_met_d("Hokusai1", breaks = c("IMF", "bvar", "ar", "bvar_qu"),
                       labels = c("IMF", "BVAR", "AR", "BVAR_QU")) +

      #scale_color_brewer(palette = "Set2") +
      geom_line() +
      geom_point() +
      ylab(ylab) +
      xlab("Forecast Horizon") +
      guides(fill = "none", colour = "none") +
      theme_uqimf() %+replace%
      theme(legend.title=element_blank(),
            legend.position = "none") +
      ggtitle(title)

  c50plot <- horizon_path_plot(cvgscores, "coverage_50", "CI Coverage - 50", facet = "error_method", title = "Central Interval Coverage - alpha = 50") +
    geom_hline(aes(yintercept = 0.5)) +
    scale_y_continuous(labels = paste0(seq(0, 100, by = 25), "%"),
                       breaks = seq(0,1, by = 0.25),
                       limits = c(0,1))

  c80plot <- horizon_path_plot(cvgscores, "coverage_80", "CI Coverage - 80", facet = "error_method", title = "Central Interval Coverage - alpha = 80") +
    geom_hline(aes(yintercept = 0.8)) +
    scale_y_continuous(labels = paste0(seq(0, 100, by = 25), "%"),
                       breaks = seq(0,1, by = 0.25),
                       limits = c(0,1))

  aeplot <- horizon_path_plot(point_scores |> .d(, error_method := "directional"), "meanae", "Absolute Error", title = "Point Forecast - Abs. Error")

  squeplot <- horizon_path_plot(point_scores |> .d(, error_method := "directional"), "meansque", "Squared Error", title = "Point Forecast - Squ. Error")

  crpsplot_rw <- horizon_path_plot(crps_scores_rw |>.d(, error_method := "directional"), "score", "CRPS by Sample", title = "CRPS - Rolling Window") +
    scale_y_continuous(limits = c(0.07,1.05))

  crpsplot_loo <- horizon_path_plot(crps_scores_loo |>.d(, error_method := "directional"), "score", "CRPS by Sample", title = "CRPS - LOO") +
    scale_y_continuous(limits = c(0.07,1.05))

  ovr_plot <- 
    (dir_wis_plot | abs_wis_plot) /
    (plot_spacer()) /
    (c50plot | c80plot) /
    (plot_spacer()) /
    (squeplot | aeplot | crpsplot_rw | crpsplot_loo) +
    plot_layout(guides = "collect", 
                heights = c(0.2, 0.05, 0.2, 0.05, 0.2)) &
      plot_annotation(tag_levels = 'I')  &
      theme(legend.position = 'bottom', 
            legend.box="vertical", legend.margin=margin())

```
\newpage
Some notes:
\begin{itemize}
\item Inflation: directional vs. absolute errors:
\begin{itemize}
\item difference small for IMF method, absolute slighly better, likely due to longer central intervals
\item AR and BVAR profit more from directional correction (upward bias)
\item for expanding window method, difference in coverage is smaller (-> structural breaks)

\end{itemize}
\item Inflation overall scores: IMF forecasts outperform others
\begin{itemize}
\item lower scores for point forecasts
\item lower WIS 
\item lower bias (compute directly?)
\end{itemize}

\item GDP Growth: more similar results for different sources
\begin{itemize}
\item lower scores at shorter horizons, more similar at larger horizons
\item IMF forecasts better only for absolute error method
\end{itemize}
\end{itemize}

\newpage

```{r gdpovr_ew, echo = FALSE, warning = TRUE, fig.dim=c(12,12)}

ovr_assessment_plot(wis_scores, scores_cvgshort, point_scores, all_scores, "ngdp_rpch")
```

# Expanding Window - Scores, by error method, Horizon and forecast source

```{r overall_assessment, fig.dim=c(12,12), echo = FALSE, warning = FALSE}

scores_cvgshort <- fread(here("scores", "ci_scores_avgcnt.csv")) |>
  setnames("model", "source", skip_absent = TRUE)

bvar_scores_cvgshort <- fread(here("scores", "bvar_ci_scores_avgcnt.csv")) |>
  setnames("model", "source", skip_absent = TRUE) |>
  .d(, source := "bvar_qu") |>
  .d(, method := "expanding window") |>
  .d(, error_method := "absolute")

bvar_scores_cvgshort <- rbind(bvar_scores_cvgshort, bvar_scores_cvgshort |> copy() |> .d(, error_method := "directional"))

point_scores <- data.table::fread(here("scores", "pointfc_scores.csv"))

wis_scores <- rbind(wis_scores, bvar_wis_scores)
scores_cvgshort <- rbind(scores_cvgshort, bvar_scores_cvgshort)

  cvgscores <- cvg_scoredat |>
    .d(method == "expanding window") |>
    .d(target == trgt) |>
    .d(, target := NULL) |>
    .d(, method := NULL) #|>

  decomp_data <- wis_scoredat |>
    .d(method == "expanding window") |>
    .d(target == trgt) |>
    .d(, target := NULL) |>
    .d(, method := NULL) #|>

  crps_scores_rw <- crps_scores |>
    .d(method == "expanding window") |>
    .d(target == trgt) |>
    .d(, target := NULL)

    if(error_m == "directional"){
      plot_name <- "WIS - Directional Errors"
    } else {
      plot_name <- "WIS - Absolute Errors"
    }

    plt <- ggplot(decomp_data |> .d(error_method == error_m),
                          aes(x = source, 
                              fill = source,
                              y = score,
                              alpha = type)) +
      geom_bar(position="stack", stat="identity") +
      facet_wrap(~ horizon,
                 labeller = as_labeller(plot_horizon_label_short(4)),
                 nrow = 1) +
      #geom_hline(yintercept = 1) +
      scale_alpha_discrete(range = c(0.35,1),
                           labels = c("Overprediction",
                                      "Underprediction",
                                      "Dispersion"),
                           name = "Components of the WIS") +
      scale_x_discrete(name = "",
                       breaks = c("IMF", "bvar", "bvar_qu", "ar"),
                       labels = c("IMF", "BVAR", "BVAR_QU", "AR")) +
      scale_y_continuous(name = "WIS", limits = c(0, maxval)) +
      scale_fill_met_d("Hokusai1") +
      #scale_fill_brewer(palette = "Set2", 
      #                  name = "Model Type")+
      xlab("") +
      #labs(fill = "Model Type") +
      theme_uqimf() %+replace%
      theme(axis.text.x = element_text(size = 8, angle = 90, hjust = .5, vjust = .5, face = "plain"),
            strip.text = element_text(size = 8)) +
      ggtitle(plot_name) +
      theme(#legend.title=element_blank(),
            legend.position = "none")

  horizon_path_plot <- function(score_data,
                                score_rule,
                                ylab = score_rule, 
                                facet = NULL,
                                title = NULL){

    path_plot <- ggplot(score_data, 
           aes(x = horizon,
               y = get(score_rule), 
               color = source)) +

      scale_color_met_d("Hokusai1", breaks = c("IMF", "bvar", "ar", "bvar_qu"),
                       labels = c("IMF", "BVAR", "AR", "BVAR_QU")) +

      #scale_color_brewer(palette = "Set2") +
      geom_line() +
      geom_point() +
      ylab(ylab) +
      xlab("Forecast Horizon") +
      guides(fill = "none", colour = "none") +
      theme_uqimf() %+replace%
      theme(legend.title=element_blank(),
            legend.position = "none") +
      ggtitle(title)

  c50plot <- horizon_path_plot(cvgscores, "coverage_50", "CI Coverage - 50", facet = "error_method", title = "Central Interval Coverage - alpha = 50") +
    geom_hline(aes(yintercept = 0.5)) +
    scale_y_continuous(labels = paste0(seq(0, 100, by = 25), "%"),
                       breaks = seq(0,1, by = 0.25),
                       limits = c(0,1))

  c80plot <- horizon_path_plot(cvgscores, "coverage_80", "CI Coverage - 80", facet = "error_method", title = "Central Interval Coverage - alpha = 80") +
    geom_hline(aes(yintercept = 0.8)) +
    scale_y_continuous(labels = paste0(seq(0, 100, by = 25), "%"),
                       breaks = seq(0,1, by = 0.25),
                       limits = c(0,1))

  aeplot <- horizon_path_plot(point_scores |> .d(, error_method := "directional"), "meanae", "Absolute Error", title = "Point Forecast - Abs. Error")

  squeplot <- horizon_path_plot(point_scores |> .d(, error_method := "directional"), "meansque", "Squared Error", title = "Point Forecast - Squ. Error")

  crpsplot_rw <- horizon_path_plot(crps_scores_rw |>.d(, error_method := "directional"), "score", "CRPS by Sample", title = "CRPS - expanding window") +
    scale_y_continuous(limits = c(0.07,1.05))

  crpsplot_loo <- horizon_path_plot(crps_scores_loo |>.d(, error_method := "directional"), "score", "CRPS by Sample", title = "CRPS - LOO") +
    scale_y_continuous(limits = c(0.07,1.05))

  ovr_plot <- 
    (dir_wis_plot | abs_wis_plot) /
    (plot_spacer()) /
    (c50plot | c80plot) /
    (plot_spacer()) /
    (squeplot | aeplot | crpsplot_rw | crpsplot_loo) +
    plot_layout(guides = "collect", 
                heights = c(0.2, 0.05, 0.2, 0.05, 0.2)) &
      plot_annotation(tag_levels = 'I')  &
      theme(legend.position = 'bottom', 
            legend.box="vertical", legend.margin=margin())

```
\newpage

```{r gdpovr, echo = FALSE, warning = TRUE, fig.dim=c(12,12)}

ovr_assessment_plot(wis_scores, scores_cvgshort, point_scores, all_scores, "ngdp_rpch")
```

scores <- fread(here("scores", "ci_scores.csv"))
scores_cvgshort <- fread(here("scores", "cvg_pooled.csv"))

cvgdat_cpin <- data.table::fread(here("scores", "cvg_pooled_directionalsymmetric.csv")) |>
  setnames("model", "source") |>
  .d(, c( "target", "error_method", "method", "source", cols), with = FALSE) |>
  melt(id.vars = c("target", "error_method", "method", "source"),
       variable.name = "pilvl",
       value.name = "coverage") |>
  .d(, pilvl := gsub("^.*?coverage_","",pilvl)) |>
  .d(, pilvl := as.numeric(pilvl)/100) |>
  .d(target == "pcpi_pch") |>
  rbind(cvgdat_cpi)

cvgdat_gdpn <- data.table::fread(here("scores", "cvg_pooled_directionalsymmetric.csv")) |>
  setnames("model", "source") |>
  .d(, c( "target", "error_method", "method", "source", cols), with = FALSE) |>
  melt(id.vars = c("target", "error_method", "method", "source"),
       variable.name = "pilvl",
       value.name = "coverage") |>
  .d(, pilvl := gsub("^.*?coverage_","",pilvl)) |>
  .d(, pilvl := as.numeric(pilvl)/100) |>
  .d(target == "ngdp_rpch") |>
  rbind(cvgdat_gdp)

\newpage
```{r, cpitable = FALSE, echo=FALSE}
#bring into long format
long_scores <- gather_scores(all_scores, "pcpi_pch")

@@
